Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-04 output/review_responses.tex

\textbf{Comment: These authors show in Fig 10-8 trial results from all over the
  rice areas of Ca, whereas the manuscript deals with trials solely
  from the RES at Briggs.This needs to be discussed and may help
  explain what is happening, as there is little or no evidence in
  McKenzie et al that, for example, the yield of M-202 is declining
  with time.}\\

\textbf{Response:} Explaination added (lines 247-269)\\

\textbf{Comment: What also worries me is that Ca farm yields (and Butte
  county yields) have been increasing since 1995 or so, at a low rate
  (about 70 kg/ha/yr) but increasing, which doesn’t agree with Fig S1
  no matter what cv you choose.}\\

\textbf{Response:} The reviewer appears to misunderstand basic
differences in the data being presented. We agree and plainly state in
the manuscript that CA farm yields have been increasing (lines 236-246). However, we do not find evidence that gains as large as 70 kg/ha/yr
are due to increasing genetic yield potential. It is entirely possible
for farm yields to increase for other reasons which, again, we outline
specifically in the manuscript (lines 241-246).\\

@@

\textbf{Response:} Table 1 and 2 have been swapped. Quality attributes
are not of direct relevance to this paper, so we have provided a
citation for where this information can be found for those
interested (line 235). We disagree that the data from Figure 2 should be
presented in table form. Data of this nature is best presented in
figures, and presenting these data twice is redundent.\\

\textbf{Comment: So in summary, I am not sure you need Fig 1, you could put the data of
  Fig 2 in Table 2 and bring it to the front along with Fig S1, and I
  don’t think you need Fig 6. But consider how you use the word
  “erosion” and discuss more thoroughly what could be causing the yield
  declines in Fig S1, and whether there is evidence it is happening also
  outside of the RES at Biggs.}\\

\textbf{Response:} Figure 1 has been moved to supplemental material,
and figure S1 has been moved to the main body. See above regarding
putting the data from figure 2 in tabular format.\\

\subsubsection{Sticky note comments}

\textbf{Comment: would "unexplained" be a better word}\\

\textbf{Response:} No - the use of inherent is intentional and conveys
a different meaning.\\

\textbf{Comment: The decline could be because the potential of the
  environment is declining for unmanageable reasons. My point is that
  your emphasis is all on the cultivars, when it has to be the
  environment deteriorating, not the cultivar. }\\

\textbf{Response: } See response above.\\

\textbf{Comment: Thats if the stats are right. Only Fig S2 really
  convinces me there is yield decline, but the cause could be
  unrelated to potential yield, as would be a manageable stress to
  which newer cvs are more resistant..}\\

\textbf{Response: } It is inappropriate for the reviewer, who above plainly
states that they do not understand the statistics, to make this
evaluation.\\

\textbf{Comment: Give sentence on location (latitudes) and area of rice in
  CA. Even a map of the rice areas , also showing location of RES}\\

\textbf{Well it went backwards for several yeasr, and then resumed
  going up but at a slower rate. Somethink happened in mid 1990s?}\\

\textbf{Response: } None needed - speculation.\\

\textbf{Comment: again it may be semantics, but the newer cvs must be
  better adapted to the deteriotating environment, thats breeding
  progress, even potential yield increase if the deterioration is part
  of the unmanageable environment (relative to the older cvs)}\\

\textbf{Response: } See response above.\\

\textbf{Comment: I cannot understand why there is no reference here in
  Introduction or in Discussion to McKenzie et al 2014 , a Chapter in
  the CSSA special publication No 33.}\\

\textbf{Response: } Added to discussion.\\

\textbf{Comment: More info on plots shape sown, and harvested please}\\

\textbf{Response: } Added to methods.\\

\textbf{Comment : Is this the best wording, as I understand year and
  cultivar in the model account for what is left after the first two
  terms. Like a bad year or a low yielding cv (maybe high quality)
  will cause a deviation of plot yield from that explained by the
  first 3 terms}\\

\textbf{Response: } This is the correct wording.\\

\textbf{Comment: is this the same as the McKay model? No, he has
  cultivar and year effetrs (relative to grand mean). And why here the
  same integer for year as for cultivar? Doesnt see quite correct the
  equation?}\\

\textbf{Response: } See response above.\\

\textbf{Comment: Progress M202 to M204 and M 205 agree with McKenzies'
  data from trials across farms and RES, with no correction for years
  after release.  But farmer data Butte County show M-206 0.4 t/ha
  over M-202. Are all their yield increases apparent because M-202 was
  eroding for longer. I guess thats what you are saying. But McKenzie
  doesnt show "erosion" of M-202 yields. Also where is M-201?}\\

\textbf{Response: } See response above. We do not have data from the
first two years of M-201, since that is before the study period.\\

\textbf{Comment: Being suspicious of any mixed liner modelling (for
  example why only 1 point for each year since release?) I find Fig S1
  as much more convincing, esp cvs like M-202, M-103, M-104, M-204,
  M-205. Why not include in the paper proper.  BUT the graphs of
  McKenzie (Fig 10-8) show now decline in M-202 yield after about
  1994. So what's going on at RES??}\\

\textbf{Response: } Again, it is wholely inappropriate for the
reviewer to be ``suspicious'' of the technique based solely on the
grounds that they do not understand it.

See response above regarding discrepencies between McKenzie et al.,
2014 and our work.\\

\textbf{Comment: agreed, but is there something which groups cvs with
  declines and ones without??}\\

\textbf{Response: } See response above.\\

\textbf{Comment: Actually its over the time between release of M 202
  and M 206, 20 years or about 0.6 t/ha.}\\

\textbf{Response: } Or exactly 560 kg/ha. \\

\textbf{Comment: Not in Fig 2 }\\

\textbf{Response: } The figure reference is correct.\\

\textbf{Comment: where are the 2 years before release for M-202 that
  you used in Fig 2. Looking at the  10 years after release when the
  two cvs can be compared (?), M-206 is no better that M-202 on this
  basis?}\\

\textbf{Response: } The data are in the figure. Yes, that is our conclusion as well.\\

\textbf{Comment: M 201 not in Fig 2. And why is there not perfect
  correspondence between Figs 2 and S2.}\\

\textbf{Response: } Figure reference updated. These figure show
different data, as indicated by the captions.\\

\textbf{Comment: Again a map would help}\\

\textbf{Response: } See response above.\\

\textbf{Comment: This is where reference ot McKay et al (2010) TAG
  122, 225-238 would be useful. They were able to pin down the
  "erosion" to increasing disease susceptibility because they had plus
  and minus fungicide. You cannot do that and I am not sure you
  separate the effect of years after release from other year related
  changes which may have been happening throughout (presumably
  negative ones) and linearly correlated with year} \\

\textbf{Response: } See response above.\\

\textbf{Comment: Could you list these supposed stresses, they may be a
  clue to the cause of erosion}\\

\textbf{Response: } See response above.\\

\textbf{Comment: But you have to argue also that breeding is
  countering these negativities such that the newest cv is yielding
  more that the older one in any given year. This is actually
  potential yield progress if the "potential"" of the E has
  deteriorated because of growing non manageable stresses like less
  sunshine, chilling, or more ozone.}\\

\textbf{Response: } See response above.\\

\textbf{Comment: see McKay et al (2010) on "erosion" due to evolving
  biotic stress organisms}\\

\textbf{Response: } See response above.\\

\textbf{Comment: This is the likely cause: a more thorough exploration
  of biotic stresses is needed : blast, nematodes, etc}\\

\textbf{Response: } As we do not have the data to explore this, this
discussion would be solely speculation. For more, see our response
above.\\

\textbf{Comment: Yes, this would be alarming....a steady decrease in
  the potential of all paddy Es, but why now after millennia of
  paddy?}\\

\textbf{Response: } As we state, more study is needed to verify and
determine the extent and cause of this (lines ).\\

\textbf{Comment: thats 1.25\% p.a. increase, almost noone is achieving
  that in any crop! Except when you jump up to hybrids, a one off
  jump. But your point is valid, yield progress is slowed by the
  "erosion" . I dont think the Figure 6,  however, is needed.}\\

\textbf{Response: } This is the rate of improvement which has been
claimed by the rice breeders in this system (lines ). The figure is
helpful to distill the major mechanisms at work here.\\

\textbf{Comment: Not in Fig 2}

\textbf{Response: } See response above.\\

\textbf{Comment: Maybe would help to mention the problems which hit in
  the mid 1990s, so that farm yield went backwards for several years,
  and then started the climb you talk about.}\\

\textbf{Response: } See response above.\\

\textbf{Comment: As shown in Butte Co in McKenxzie et al 2014, another
  reason for a map!}\\

\textbf{Response:}\\

\textbf{Comment: Again, a special period of instabiliy from early
  1990s for a decade or so. Why} \\

\textbf{Response: } See response above.\\

\textbf{Comment: Not sure you need Fig 6 or the last paragraph. The
  issue is we need to know why erosion is occuring , and only it seems
  in some cultivars (see Fig S2?)}\\

\textbf{Response: } See response above.\\

\textbf{Comment: This is only possible if you account also for all
  environmental changes (climate plus agronomy)....you need to be
  running a rice simulation model continuously. Note climate is more
  than just T, also radiation, ratio diffuse to total, ozone, CO2 ),
  and with paddy in Ca there is also water quality and
  temperature. And agronomy is also gradual soil changes, fallow may
  not be enough.}\\

\textbf{Response: } See response above.\\

\textbf{Comment: Explain better what you mean}\\

\textbf{Response: } Done.\\

\textbf{Comment: Improved management can contribute to increased
  potential yield, if its "new" agronomic technology.}\\

\textbf{Response: } See response above, as well as the reviewer's own
definition of yield potential.\\

\textbf{Response: } See response above.\\

\textbf{Comment: are some dot black and some grey or is it my eyes?}\\

\textbf{Response: } See response above.\\

@@

\textbf{Comment: L53: “Trials” instead of “Trails”}\\

\textbf{Response:} Corrected.\\

\textbf{Comment: L120: (Fig. 2)}\\

\textbf{Comment: L141: Something is wrong with “…since release is between…”}\\

### 2018-03-04 output/breed_paper_revised.Rnw

Experimental plots at the RES are managed similar to University of
California Cooperative Extension prescribed best management practice
for land preparation, fertility rates, and pest and disease
control~\citep{ucce}. As is typical for water-seeded systems, the
plots were planted by pre-germinating rice seed prior to
direct-seeding into pre-flooded fields. Plots were either 14.0 or 18.5
m\tsups{2}, and were harvested after physiological maturity using a
small-plot combine. All yields were converted to 14\% MC prior to
reporting.

@@

After correcting for yield decline over time, some cultivars were higher
yielding compared to the overall average (as estimated via the random
intercept term in the model). The marginal estimate for the
performance of a cultivar averaged over the uncertainty in the other
parameters in the model (i.e., the random effect) shows three
cultivars with higher than average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),colla=', ')}
(Figs. \ref{fig:first_yrs} \& \ref{fig:ids_comp}).

@@

\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg\tsups{-1} ha\tsups{-1}
year\tsups{-1}. Yet the CI estimated from these data using the best
model suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg\tsups{-1} ha\tsups{-1} year\tsups{-1}. There are several
explanations for how this could be. 1) \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. 2) These yield decreases over time
were not taken into account possibly because \citet{mckenzie2014}
concluded that there was no trend in the yield of the major cultivars
over time. However this conclusion is based on a flawed interpretation
of the results of their statistical test. Failing to reject the
hypothesis that the trend in yield equals zero does not equate to the
confirmation of no trend. Instead, failing to reject the hypothesis
could suggest the analysis lacks the statistical power to detect the trend. 3) It
is not surprising that no trend in yield over time was detected given
that \citet{mckenzie2014} used data from several locations, yet
analyzed those data using a simple t-test. By using this simple
analysis, \citet{mckenzie2014} ignore systematic errors related to
differences between observations from different locations, including
different management systems, growing environments, and crop
timing. By contrast, this investigation presented here used a single,
well managed system and analyzed the data using a mixed effects model
which fully accounts for non-independence between similar
observations. In doing this, the analysis makes fewer assumptions and
more fully captures the structural relationships in the data. Hence,
the estimates presented here are likely more reliable.

>>>>

Experimental plots at the RES are managed similar to University of
California Cooperative Extension prescribed best management practice
for land preparation, fertility rates, and pest and disease
control~\citep{ucce}. As is typical for water-seeded systems, the
plots were planted by pre-germinating rice seed prior to
direct-seeding into pre-flooded fields. Plots were rectangular and either 14.0 or 18.5
m\tsups{2}, and were harvested after physiological maturity using a
small-plot combine. All yields were converted to 14\% MC prior to
reporting.

@@

After correcting for yield decline over time, some cultivars were higher
yielding compared to the overall average (as estimated via the random
intercept term in the model). The marginal estimate for the
performance of a cultivar averaged over the uncertainty in the other
parameters in the model (i.e., the random effect) shows three
cultivars with higher than average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),colla=', ')}
(Fig. \ref{fig:ids_comp}).

@@

\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg\tsups{-1} ha\tsups{-1}
year\tsups{-1}. Yet the CI estimated from these data using the best
model suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg\tsups{-1} ha\tsups{-1} year\tsups{-1}. There are several
explanations for this discrepancy in conclusions. First, \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. Second, these yield decreases over time
were not taken into account possibly because \citet{mckenzie2014}
concluded that there was no trend in the yield of the major cultivars
over time. However this conclusion is based on a flawed interpretation
of the results of their statistical test. Failing to reject the
hypothesis that the trend in yield equals zero does not equate to the
confirmation of no trend. Instead, failing to reject the hypothesis
could suggest the analysis lacks the statistical power to detect the
trend. Third, it
is not surprising that no trend in yield over time was detected given
that \citet{mckenzie2014} used data from several locations, yet
analyzed those data using a simple t-test. By using this simple
analysis, \citet{mckenzie2014} ignore systematic errors related to
differences between observations from different locations, including
different management systems, growing environments, and crop
timing. By contrast, the investigation presented here used a single,
well managed system and analyzed the data using a mixed effects model
which fully accounts for non-independence between similar
observations. In doing this, the analysis makes fewer assumptions and
more fully captures the structural relationships in the data. Hence,
the estimates presented here are likely more reliable.
